---
title: "Mini Project 2"
author: "Mae Rosenstein, Chelsey Chen, Talia Seshaiah"
date: "March 22, 2018"
output:
  html_document:
    code_folding: show
---

The data represented in these visualizations originate from the Federal Election Commission. The data is focused around the 2011 - 2013 election cycle. The candidates are those who have registered with the Federal Election Commission or are present on an official ballot list from a given state's elections office. The contributions to candidates are ones that have been received from itemized committees, PAC contributions to a party committee, party transfers between state committees, and party transfers from a national committee to a state committee. The contributions highlighted in these visualizations are only independent expenditures advocating for election of the specified candidate.
```{r, include=FALSE, message = FALSE, warning = FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(fiftystater)
library(mapproj)
library(RColorBrewer)
```
## Data Wrangling 
#####We first filtered the data to only include "24E" which are only independent expenditures advocating for the election of a candidate. We then filtered only looked at positive transaction amount, therefore eliminating out any donations that were returned. We joined the candidates dataset with contributions dataset by candidate's id number which was a variable present in both datasets. We selected variables that we were particularly interested in looking at: candidate's party affiliation, transaction amount, transaction date, candidate name, and the source of the contributions (entity type) which we filtered to be just organizations.We then mutated the transaction date from numeric numbers to strings, and renamed the candidate party affiliation names to display the full names rather than the abbreviation. 

```{r}
contri_candi <- contributions %>% 
  filter(transaction_type == "24E",
         transaction_amt >= 0,
         entity_type != "") %>%
  left_join(candidates, by = "cand_id") %>%
  filter(cand_party_affiliation %in% c('REP','DEM','IND','LIB')) %>%
  select(cand_party_affiliation, transaction_amt, transaction_dt, cand_name, entity_type,state) %>%
  mutate(date = as.Date(transaction_dt, "%m%d%Y"),
         year = substr(date,0,4)) %>%
  select(-transaction_dt) %>%
  filter(date <= '2012-12-31',
         date >= '2011-01-01') %>%
  mutate(cand_party_affiliation = ifelse(cand_party_affiliation == 'REP', 'Republican', cand_party_affiliation),
         cand_party_affiliation = ifelse(cand_party_affiliation == 'DEM', 'Democratic', cand_party_affiliation),
         cand_party_affiliation = ifelse(cand_party_affiliation == 'LIB', 'Libertarian', cand_party_affiliation),
         cand_party_affiliation = ifelse(cand_party_affiliation == 'IND', 'Independent', cand_party_affiliation)) %>%
  group_by(cand_name,cand_party_affiliation,date,year,entity_type,state) %>%
  summarize(total_amt = sum(transaction_amt))
```

## Data Visualizations
### Figure 1: The contribution amount given to the republican party, democratic party, independent party, and libertarian party from January 1, 2011 to December 31, 2012.
```{r}
<<<<<<< HEAD
graph_fun <- function(entity,colour){
=======
# Write a function to plot the graph given a specific source of contribution and a color
graph_fun <- function(entity, colour){
>>>>>>> 64d180c06f5d77ca34fef24403c79e2dd612e349
  contri_candi %>% 
  filter(entity_type == entity) %>% 
  ggplot(aes(x=date, y= total_amt)) +
  geom_line(color = colour) +
  theme_bw() +
  labs(x = NULL,
       y = "Total Amount of Contributions to Candidates (US dollar)") +
  scale_y_continuous(breaks = c(0,2000000,4000000,6000000,8000000),
                     labels = c("0","2M", "4M", "6M","8M")) +
  theme(axis.text.x = element_text(angle = 45, hjust =1),
        panel.grid.major = element_line(colour = "grey94"),
        panel.grid.minor = element_line(colour = "white")) +
  facet_grid(~cand_party_affiliation)
}
<<<<<<< HEAD
my_plot1 <- graph_fun("ORG","hotpink4")
my_plot1
=======

# Call the graphing function for contributions made by orgnizations
my_plot1 <- graph_fun("ORG","hotpink4")
my_plot1
# # Call the graphing function for contributions made by individuals
>>>>>>> 64d180c06f5d77ca34fef24403c79e2dd612e349
my_plot2 <- graph_fun("IND","navyblue")
my_plot2
```
This visualization displays the total amount of contributions by party affiliation. The data was filtered to span from January 1, 2011 to January 1, 2013. The contributions included in the visualization are from various types of contriubtors ranging from large organizations to individuals. The parties highlighted in the visualization are democratic party, republican party, libertarian party, and independent. The republican and democratic parties received the most contributions, however the republican party received the most contributions as seen in the visualization. These results are not surprising due to the fact the the democratic party and republican party are the primary political parties. 


### Figure 2: Contributions made by different states
```{r}
# Read 52 state codes into a local dataframe [includes DC (Washington D.C. and PR (Puerto Rico)]
st.codes<-data.frame(abb=as.factor(c("AK", "AL", "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA",
                                         "HI", "IA", "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME",
                                         "MI", "MN", "MO", "MS",  "MT", "NC", "ND", "NE", "NH", "NJ", "NM",
                                         "NV", "NY", "OH", "OK", "OR", "PA", "PR", "RI", "SC", "SD", "TN",
                                         "TX", "UT", "VA", "VT", "WA", "WI", "WV", "WY")),
                      state=as.factor(c("alaska","alabama","arkansas","arizona","california","colorado",
                                       "connecticut","district of columbia","delaware","florida","georgia",
                                       "hawaii","iowa","idaho","illinois","indiana","kansas","kentucky",
                                       "louisiana","massachusetts","maryland","maine","michigan","minnesota",
                                       "missouri","mississippi","montana","north carolina","north dakota",
                                       "nebraska","new hampshire","new jersey","new mexico","nevada",
                                       "new york","ohio","oklahoma","oregon","pennsylvania","puerto rico",
                                       "rhode island","south carolina","south dakota","tennessee","texas",
                                       "utah","virginia","vermont","washington","wisconsin",
                                       "west virginia","wyoming"))
                     )

# Join the created st.codes dataframe with our dataframe by the state name
contri_candi <- contri_candi %>%
  right_join(st.codes, state = state) %>%
  select(-abb)

# Create a summary dataframe calculating total contributions for a given party affiliation,year,source,and state
state.summary <- contri_candi %>%
  group_by(cand_party_affiliation, year, entity_type, state) %>%
  summarize(total = sum(total_amt)) %>%
  filter(!is.na(total))

# Create a custom color palette (diverging colors)
myPalette <- colorRampPalette(rev(brewer.pal(9, "RdGy")))

# Create a map showing contributions to Republican candidates made by orgnizations in 2012 by different states
my_plot3 <- state.summary %>% 
  filter(entity_type == "ORG",
         cand_party_affiliation == "Republican",
         year == 2012) %>%
  ggplot(aes(map_id = full)) +
  geom_map(aes(fill = total), map = fifty_states) +
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradientn(colors = myPalette(4:9),
                       breaks=c(1000,50000000),
                       labels=c("1K","50M"),
                       limits=c(1000,50000000),
                       guide = guide_colorbar(title = "Amount of Contributions")) +
  labs(x = "", y = "") +
  theme(panel.background = element_blank())

my_plot3
```


Because the republican party and democratic party revealed to receive the most contributions, the following visualizations display the distribution of contributions from every state to each party.The contributions highlighted in this visualzation are only contributions from organizations. In the data visualzation of the republican party, Maryland clearly recived the most contributions. 

Link to GitHub repository: https://github.com/mrosenstein/sds192-mp2.git
