---
title: "Mini Project 2"
author: "Mae Rosenstein, Chelsey Chen, Talia Seshaiah"
date: "March 22, 2018"
output: html_document:
    code_folding: show
---

We looked at contribution amount to each candidates and how that differs by candidate's party affiliation. We also looked at how contributions change over time by using time series data.

Data wrangling:
1) We first filter out
2) Join the candidates dataset with contribution dataset by candidate's id number;
3) We only looked at transaction type 24E, which is independent expenditure advocating election of candidate. 
4) We only looked at positive transaction amount, therefore filtering out donations that are returned.
4) We only select variables that we are interestd in looking at: candidate's party affiliation, transaction amount, transaction date, candidate name, and the source of the contributions (the entity type (CAN, CCM, COM, ORG, PAC, PTY));
5) We mutate the transaction date from numeric numbers to strings;
```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```

```{r}
library(tidyverse)
glimpse(house_elections)
glimpse(candidates)
glimpse(committees)
glimpse(contributions)
```


```{r}

contri_candi <- contributions %>% 
  filter(transaction_type == "24E",
         transaction_amt >= 0,
         entity_type != "") %>%
  left_join(candidates, by = "cand_id") %>%
  filter(cand_party_affiliation %in% c('REP','DEM','IND','LIB')) %>%
  select(cand_party_affiliation, transaction_amt, transaction_dt, cand_name, entity_type) %>%
  mutate(date = as.Date(transaction_dt, "%m%d%Y"),
         year = substr(date,0,4)) %>%
  select(-transaction_dt) %>%
  mutate(cand_party_affiliation = ifelse(cand_party_affiliation == 'REP', 'R', cand_party_affiliation),
         cand_party_affiliation = ifelse(cand_party_affiliation == 'DEM', 'D', cand_party_affiliation)) %>%
  group_by(cand_name,cand_party_affiliation,date,year,entity_type) %>%
  summarize(total_amt = sum(transaction_amt))
           
<<<<<<< HEAD
contri_candi <- contri_candi %>%
  mutate(cand_party_affiliation = case_when(
  cand_party_affiliation == "REP" ~ "R",
  cand_party_affiliation == "DEM" ~ "D")
 )

=======
glimpse(contri_candi)
>>>>>>> c44b7b9452a22decb3999a02a0d053350f1d7fa5
```

```{r}
my_plot <- contri_candi %>% ggplot(aes(x=date, y= total_amt,color=cand_party_affiliation)) +
  geom_point() +
  theme_bw()
  
my_plot
```
### Contributions to candidates with different party affiliations
```{r}
my_plot2 <- contri_candi %>% ggplot(aes(x=date, y= total_amt)) +
  #geom_point(size =0.5) +
  geom_line() +
  theme_bw() +
  labs(x = NULL,
       y = "Total Expenditures to Candidates") +
  scale_y_continuous(breaks = c(0,2000000,4000000,6000000,8000000),
                     labels = c("0","2M", "4M", "6M","8M")) +
  theme(axis.text.x = element_text(angle = 45, hjust =1)) +
  facet_grid(~cand_party_affiliation)
  
my_plot2
```
### Contributions made by different entities
```{r}
my_plot3 <- contri_candi %>% 
  filter(cand_party_affiliation == "REP") %>%
  ggplot(aes(x=date, y= total_amt)) +
  geom_line()+
  theme_bw() +
  labs(x = NULL,
       y = "Total Expenditures to Candidates") +
  scale_y_continuous(breaks = c(0,2000000,4000000,6000000,8000000),
                     labels = c("0","2M", "4M", "6M","8M")) +
  #scale_x_date(breaks = c(2011-01-01,2012-01-01,2013-01-01),
  #                   labels = c("2011","2012","2013")) +
  theme(axis.text.x = element_text(angle = 45, hjust =1)) +
  facet_grid(~entity_type)

my_plot3
```

```{r}
summary <- contri_candi %>% 
  group_by(cand_name, cand_party_affiliation, entity_type, year) %>% 
  summarise(total = sum(total_amt))
summary

my_plot4 <- summary %>% 
  filter(cand_party_affiliation == "REP") %>%
  ggplot(aes(x=year, y= total)) +
  geom_bar()+
  theme_bw() +
  labs(x = NULL,
       y = "Total Expenditures to Candidates") +
  scale_y_continuous(breaks = c(0,2000000,4000000,6000000,8000000),
                     labels = c("0","2M", "4M", "6M","8M")) +
  #scale_x_date(breaks = c(2011-01-01,2012-01-01,2013-01-01),
  #                   labels = c("2011","2012","2013")) +
  #theme(axis.text.x = element_text(angle = 45, hjust =1)) +
  facet_grid(~entity_type)

my_plot4
```
The data represented in these visualizations originate from the Federal Election Commission. The data is focused around the 2011 - 2012 election cycle. The candidates are those who have registered with the Federal Election Commission or are present on an official ballot list from a given state's elections office. The contributions to candidates are ones that have been received from itemized committees, PAC contributions to a party committee, party transfers between state committees, and party transfers from a national committee to a state committee. The contributions highlighted in these visualizations are only independent expenditures advocating for election of the specified candidate. 

This visualization displays the total amount of contributions by party affiliation. The contributions included in the visualization are from various types of contriubtors ranging from large organizations to individuals. The republican and democratic parties received the most contributions, however the republican party received the most contributions as seen in the visualization. 

Because the republican party revealed to receive the most contributions, the second visualization displays which type of contributors contributed the most to the republican party. As seen in the visualization, organizations contributed the most to the republican party. 

Key for Candidate Party Affiliation: 
DEM = Democratic Party
DFL = Democratic-Farmer-Labor
IND = Independent 
LIB = Libertarian Party
NNE = None
OTH = Other 
REP = Republican Party 

Key for Contribution Type to Candidates:  
CAN = Candidate
CCM = Candidate Committee
COM = Committee
IND = Individual (a person)
ORG = Organization (not a committee and not a person)
PAC = Political Action Committee
PTY = Party Organization